Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-10-01 review/response.Rmd

\textcolor{red}{The data set is first introduced in the subsection Simulation setup under Diagnosing an optimiser. To avoid confusion, we change the wording from "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...".}

-   "While the previous few examples have looked at the space of 1D bases ... 2D bases"

    -   I don't think the terms "1D bases" and "2D bases" are standard. Perhaps (orthonormal) "1-frames" and "2-frames"?

\textcolor{red}{You are right that we could use frame instead of basis, and both are used conventionally. We have elected to stay with the use of basis, and have  some sentences in the section titled "Projection pursuit guided tour" to explain the terminology.}

### 2021-10-01 paper/zhang-cook-laa-langrene-menendez.Rmd
last tweaks

# Projection pursuit guided tour {#tour}

Projection pursuit guided tour combines two different methods (projection pursuit and guided tour) to explore interesting features in a high-dimensional space. Projection pursuit, coined by @friedman1974projection, detects interesting structures (e.g. clustering, outliers and skewness) in multivariate data via low-dimensional projections. Guided tour [@cook1995grand] is one variation of a broader class of data visualisation methods, tour [@buja2005computational], which displays high-dimensional data through a series of animated projections.

@@

As a general visualisation method, tour produces animations of high-dimensional data via rotations of low-dimensional planes. There are different versions depending on how the high-dimensional space is investigated: grand tour [@cook2008grand] selects the planes randomly to provide a general overview; manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection. Guided tour, the main interest of this paper, chooses the planes with the aid of projection pursuit, to gradually reveal the most interesting projection. Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs a geodesic interpolation between these planes to form a tour path. Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:tour-path}','\\@ref(fig:tour-path)'))` shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation and the white frames are interpolations between the blue frames. Mathematical details of the geodesic interpolation can be found in @buja2005computational. Notice that the term *frame* used in Buja's paper is conventionally called *basis*, which is used in this paper and the associated software. The aforementioned tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].

>>>>

# Projection pursuit guided tour

A projection pursuit guided tour combines two different methods (projection pursuit and guided tour) to explore interesting features in a high-dimensional space. Projection pursuit, coined by @friedman1974projection, detects interesting structures (e.g. clustering, outliers and skewness) in multivariate data via low-dimensional projections. Guided tour [@cook1995grand] is one variation of a broader class of data visualisation methods, tour [@buja2005computational], which displays high-dimensional data through a series of animated projections.

@@

As a general visualisation method, tour produces animations of high-dimensional data via rotations of low-dimensional planes. There are different versions depending on how the high-dimensional space is investigated: grand tour [@cook2008grand] selects the planes randomly to provide a general overview; manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection. Guided tour, the main interest of this paper, chooses the planes with the aid of projection pursuit, to gradually reveal the most interesting projection. Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs a geodesic interpolation between these planes to form a tour path. Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:tour-path}','\\@ref(fig:tour-path)'))` shows a sketch of the tour path where the blue squares represent planes (targets) selected by the projection pursuit optimisation and the white squares represent planes in the geodesic interpolation between targets. Mathematical details of the geodesic interpolation can be found in @buja2005computational. (Note that the term *frame* used in Buja's paper to refer to a particular set of orthonormal vectors defining a plane. This is also conventionally referred to as a basis, which is used in this paper and the associated software.) The aforementioned tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].

### 2021-10-01 paper/zhang-cook-laa-langrene-menendez.tex
last tweaks

\hypertarget{tour}{%
\section{Projection pursuit guided tour}\label{tour}}

Projection pursuit guided tour combines two different methods
(projection pursuit and guided tour) to explore interesting features in
a high-dimensional space. Projection pursuit, coined by
\citet{friedman1974projection}, detects interesting structures
(e.g.~clustering, outliers and skewness) in multivariate data via
low-dimensional projections. Guided tour \citep{cook1995grand} is one
variation of a broader class of data visualisation methods, tour
\citep{buja2005computational}, which displays high-dimensional data
through a series of animated projections.

@@

As a general visualisation method, tour produces animations of
high-dimensional data via rotations of low-dimensional planes. There are
different versions depending on how the high-dimensional space is
investigated: grand tour \citep{cook2008grand} selects the planes
randomly to provide a general overview; manual tour
\citep{cook1997manual} gradually phases in and out one variable to
understand the contribution of that variable in the projection. Guided
tour, the main interest of this paper, chooses the planes with the aid
of projection pursuit, to gradually reveal the most interesting
projection. Given a random start, projection pursuit iteratively finds
bases with higher index values and the guided tour constructs a geodesic
interpolation between these planes to form a tour path. Figure
\ref{fig:tour-path} shows a sketch of the tour path where the blue
frames are produced by the projection pursuit optimisation and the white
frames are interpolations between the blue frames. Mathematical details
of the geodesic interpolation can be found in
\citet{buja2005computational}. Notice that the term \emph{frame} used in
Buja's paper is conventionally called \emph{basis}, which is used in
this paper and the associated software. The aforementioned tour method
has been implemented in the R package \CRANpkg{tourr} \citep{tourr}.

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/tour-path}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-search}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-interp}

@@

{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-pca}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-pca-animated-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-tour-1}

@@

\caption{Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}. The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo. The full animation can be seen in the html version of the paper.}\label{fig:toy-tour}
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-torus-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/interrupt}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/polish}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/noisy-better-geo}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/kol-result}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/flip-sign}

@@

>>>>

\hypertarget{projection-pursuit-guided-tour}{%
\section{Projection pursuit guided
tour}\label{projection-pursuit-guided-tour}}

A projection pursuit guided tour combines two different methods
(projection pursuit and guided tour) to explore interesting features in
a high-dimensional space. Projection pursuit, coined by
\citet{friedman1974projection}, detects interesting structures
(e.g.~clustering, outliers and skewness) in multivariate data via
low-dimensional projections. Guided tour \citep{cook1995grand} is one
variation of a broader class of data visualisation methods, tour
\citep{buja2005computational}, which displays high-dimensional data
through a series of animated projections.

@@

As a general visualisation method, tour produces animations of
high-dimensional data via rotations of low-dimensional planes. There are
different versions depending on how the high-dimensional space is
investigated: grand tour \citep{cook2008grand} selects the planes
randomly to provide a general overview; manual tour
\citep{cook1997manual} gradually phases in and out one variable to
understand the contribution of that variable in the projection. Guided
tour, the main interest of this paper, chooses the planes with the aid
of projection pursuit, to gradually reveal the most interesting
projection. Given a random start, projection pursuit iteratively finds
bases with higher index values and the guided tour constructs a geodesic
interpolation between these planes to form a tour path. Figure
\ref{fig:tour-path} shows a sketch of the tour path where the blue
squares represent planes (targets) selected by the projection pursuit
optimisation and the white squares represent planes in the geodesic
interpolation between targets. Mathematical details of the geodesic
interpolation can be found in \citet{buja2005computational}. (Note that
the term \emph{frame} used in Buja's paper to refer to a particular set
of orthonormal vectors defining a plane. This is also conventionally
referred to as a basis, which is used in this paper and the associated
software.) The aforementioned tour method has been implemented in the R
package \CRANpkg{tourr} \citep{tourr}.

@@

{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/tour-path}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-search}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-interp}

@@

{\centering \includegraphics[width=0.7\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-pca}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-pca-animated-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-tour-1}

@@

\caption{Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}. The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo. The full animation can be seen in the html version of the paper.}(\#fig:toy-tour)
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-torus-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/interrupt}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/polish}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/noisy-better-geo}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/kol-result}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/flip-sign}

@@

\begin{verbatim}
\end{verbatim}
